\subsection{Sanctum}
\label{sec:related_sanctum}

Sanctum~\cite{costan2015sanctum} introduced a straightforward software/hardware
co-design that yields the same resilience against software attacks as SGX, and
adds protection against memory access pattern leaks, such as page fault
monitoring attacks and cache timing attacks.

Sanctum uses a conceptually simple cache partitioning scheme, where a
computer's DRAM is split into equally-sized continuous DRAM regions, and each
DRAM region uses distinct sets in the shared last-level cache (LLC). Each DRAM
region is allocated to exactly one container, so containers are isolated in
both DRAM and in the LLC. Containers are isolated in the other caches by
flushing on context switches.

Like in XOM, Aegis and Bastion, the hypervisor, OS, and the application
software outside containers are conceptually considered to belong to a separate
container. Containers are protected from the untrusted outside software by the
same measures that isolate containers from each other.

Sanctum relies on a trusted security monitor, which is the first piece of
firmware executed by the processor, and has the same security properties as
Aegis' security kernel. The monitor is measured by bootstrap code in the
processor's ROM, and its cryptographic hash is included in the software
attestation measurement. The monitor verifies the operating system's resource
allocation decisions. For example it ensures that no DRAM region is ever
accessible to two different containers.

Each Sanctum container manages its own page tables, mapping its DRAM regions,
and handles its own page faults. It follows that an malicious OS cannot learn
a container's page faults addresses. Sanctum's hardware modifications work in
conjunction with the security monitor to make sure that a container's page
tables only reference memory inside the container's DRAM regions.

The Sanctum design focuses completely on software attacks, and does not offer
protection from any physical attack. The authors expect Sanctum's hardware
modifications to be combined with the physical attack protections in Aegis or
Ascend.
