\section{Computer Architecture Background}
\label{sec:architecture_background}

This section attempts to summarize the general architectural principles behind
Intel's most popular computer processors, as well as the peculiarities needed
to reason about the security properties of a system running on these
processors. Unless specified otherwise, the information here is summarized from
Intel's \textit{Software Development Manual}~(SDM)~\cite{intel2015sdm}.

Analyzing the security of a software system requires understanding the
interactions between all the parts of the software's execution environment, so
this section is quite long. We do refrain from introducing any security
concepts here, so readers familiar with x86's intricacies can safely skip this
section and refer back to it when necessary.

We use the terms \textit{Intel processor} or \textit{Intel CPU} to refer to the
server and desktop versions of Intel's Core line-up. In the interest of space
and mental sanity, we ignore Intel's other processors, such as the embedded
line of Atom CPUs, or the failed Itanium line. Consequently, the terms
\textit{Intel computers} and \textit{Intel systems} refers to computer systems
built around Intel's Core processors.

In this paper, the term \textit{Intel architecture} refers to the x86
architecture described in Intel's SDM. The x86 architecture is overly complex,
mostly due to the need to support executing legacy software dating back to 1990
directly on the CPU, without the overhead of software interpretation. We only
cover the parts of the architecture visible to modern 64-bit software, also
in the interest of space and mental sanity.

The 64-bit version of the x86 archicture, covered in this section, was actually
invented by Advanced Micro Devices (AMD), and is also known as AMD64,
\texttt{x86\_64}, and x64. The term ``Intel architecture'' highlights our
interest in the architecture's implementation in Intel's chips, and our desire
to understand the mindsets of Intel SGX's designers.


\input{contents/architecture/overview.tex}
\input{contents/architecture/resources.tex}
\input{contents/architecture/rings.tex}
\input{contents/architecture/addresses.tex}
\input{contents/architecture/paging.tex}
\input{contents/architecture/registers.tex}
\input{contents/architecture/segments.tex}
\input{contents/architecture/faults.tex}
\input{contents/architecture/computer.tex}
\input{contents/architecture/out_of_order.tex}
\input{contents/architecture/caching.tex}
\input{contents/architecture/interrupts.tex}
\input{contents/architecture/booting.tex}
\input{contents/architecture/microcode.tex}
