\subsection{Interrupts}
\label{sec:interrupts}

Peripherals use \textit{interrupts} to signal the occurrence of an event that
must be handled by system software. For example, a keyboard issues interrupts
when a key is pressed or depressed. System software also relies on interrupts
to implement preemptive multi-threading. This section presents the details
needed to understand the concerns and tools that interrupts bring to the SGX
implementation.

% Advanced Programmable Interrupt Controller (APIC): SDM S 10
% Message Signalled Interrupts (MSI): SDM S 10.11

Peripherals use bus-specific protocols to signal interrupts. For example, PCIe
relies on \textit{Message Signaled Interrupts} (MSI), which are memory writes
issued to specially designed memory addresses. The bus-specific interrupt
signals are received by the \textit{I/O Advanced Programmable Interrupt
Controller} (IOAPIC) in the PCH, shown in Figure~\ref{fig:motherboard}.

% Local APIC ID: SDM S 10.4.6
% Extended xAPIC (x2APIC): SDM S 10.12

The IOAPIC routes interrupt signals to one or more \textit{Local Advanced
Programmable Interrupt Controllers} (LAPICs). As shown in
Figure~\ref{fig:cpu_die}, each logical CPU has a LAPIC that can receive
interrupt signals from the IOAPIC. The IOAPIC routing process assigns each
interrupt to an 8-bit \textit{interrupt vector}, used to identify the interrupt
sources, and a 32-bit \textit{APIC ID} used to identify the LAPIC that receives
the interrupt.

% Handling Interrupts: SDM S 10.8

Each LAPIC uses a 256-bit \textit{Interrupt Request Register} (IRR) to track
the un-serviced interrupts that it has received, based on the interrupt vector
number. When the corresponding logical processor is available, the LAPIC copies
the highest-priority un-serviced interrupt vector to the
\textit{In-Service Register} (ISR), and invokes the logical processor's
interrupt handling process.

% Issuing Interprocessor Interrupts: SDM S 10.6
% Interrupt Command Register (ICR): SDM S 10.6.1

At the architectural level, interrupt handling reuses many of the mechanisms in
fault handling (\S~\ref{sec:faults}). The interrupt vector number in the
LAPIC's ISR is used to locate an interrupt handler in the IDT, and the handler
is invoked, possibly after a privilege switch is performed. The interrupt
handler does the processing that the device requires, and then writes the
LAPIC's \textit{End Of Interrupt} (EOI) register to signal the fact that it has
completed handling the interrupt.

Interrupts are treated like faults, so interrupt handlers have full control
over the execution environment of the application being interrupted. This is
used to implement pre-emptive multi-threading, which relies on a clock device
that generates interrupts periodically, and on an interrupt handler that
performs context switches. While useful for multi-threading, this aspect of the
architecture also implies that the SGX design must include changes to the
interrupt handling process, in order to protect an enclave's execution state
from an interrupt handler that is invoked while the enclave's code is
executing.

System software can cause an interrupt on any logical processor by writing the
target processor's APIC ID into the \textit{Interrupt Command Register} (ICR)
of the LAPIC associated with the logical processor that the software is runing
on. These interrupts, called \textit{Inter-Processor Interrupts} (IPI), are
needed to implement TLB shoot-downs (\S~\ref{sec:tlbs}). The IPI mechanism can
also be used to send SMI interrupts, which opens up an avenue for a malicious
kernel or hypervisor to mount a SMM mode (\S~\ref{sec:rings}) attack.
